---
title: "Week4 - Assignment"
author: "M. Onimus"
date: "9/27/2020"
output: 
  html_document:
      code_folding: hide
      toc: yes
      toc_depth: 3
      toc_float:
        collapse: yes
        smooth_scrolling: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


```{r libs, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(gtsummary)
library(gt)
library(gtsummary)
library(haven)
library(readxl)
library(rlang)
library(rstatix)
```

# Read in Data 

```{r readData}

school <- read_sav("assignments/BasicTests_School.sav") %>% 
  mutate(School = as.factor(School))


```


# Question 1

Chichi and David are still working on their study of what children are eating for lunch at school. In the BasicTests School.sav file are 8 random samples of meals from 3 schools (hence 24 observations total). The data represents fat content in grams from meals served at the schools. The two are interested in whether there is a difference in fat content of the lunches between schools. Please test that hypothesis using an ANOVA.

## Question/Answer 1a

What are the null and alternative hypotheses? 

**H~0~: $\mu_1=\mu_2, \mu_1=\mu_3, \mu_2=\mu_3$**

**H~A~: $\mu_i \neq \mu_m$ for some $i\neq m$**

## Question/Answer 1b

What is the average fat content in the lunches of each of the schools?

```{r 1b}

sumSchool <- school %>% 
  group_by(School) %>% 
  summarize(mean = mean(FatContent))

sumSchool %>% 
  gt() %>% 
  tab_header(
    title = "Means of FatContent by School"
  )



```

**The mean fat content of school 1 is `r sumSchool$mean[[1]]`, school 2 is `r sumSchool$mean[[2]]`, and school 3 is `r sumSchool$mean[[3]]`.**

## Question/Answer 1c

What is the test statistic? What are the degrees of freedom?

```{r 1c}



a1 <- aov(FatContent ~ School, data = school)

aSum <- summary(a1) # it is tough to pull apart the anova model, easier with the summary


pander::pander(anova(a1))




```

**The F-value test statistics is `r aSum[[1]][["F value"]][[1]]` and the degrees of freedom are `r aSum[[1]][["Df"]][[1]]`.**

## Question/Answer 1d

What is the p-value associated with that test statistic? 

**The associated $p$-value with the test statistic is `r aSum[[1]][["Pr(>F)"]][[1]]`.**

## Question/Answer 1e

What can you say about the model assumption of homogeneous variances between groups? Please include any relevant plots or descriptive/inferential statistics as appropriate to support your discussion.

```{r 1f}

levene <- levene_test(FatContent ~ School, data = school)

ggplot(school, aes(x = School, y = FatContent)) +
  geom_boxplot(aes(fill = School), alpha = 0.3) +
  geom_jitter() +
  theme_classic()
  


```

**The Levene F-value test statistic from the Levene test was `r levene$statistic` with a corresponding $p$-value of `r levene$p`.  Indicating we can assume consistent variance.  This is also demostrated by the boxplot showing the distrubtion for each school is approximately the same.**

**NOTE: these results are different from what is produced in SAS and I am not sure why yet.**

## Question/Answer 1f

Are post-hoc tests appropriate? If so, please report on which schools differ in fat
content, using a Bonferroni correction. 

```{r 1g}

pwt <- school %>% 
  pairwise_t_test(FatContent ~ School, p.adjust.method = "bonferroni")

pwt

```

**The $p$-value of `r aSum[[1]][["Pr(>F)"]][[1]]` from the ANOVA indicated that no group was significantly different; an post-hoc analysis may not be needed as we have answered the main question already. As a practice, I have completed a pairwise t test regardless.  As the ANOVA indicated, there is no significant differences between any school in terms of fat content.**


## Question/Answer 1g

Write up your interpretation of the results of your analysis.

****

