---
title: "Week3-Basic Tests"
author: "Matthew Onimus"
date: "9/22/2020"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapse: true
      smooth_scrolling: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libs, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(gtsummary)
library(gt)
library(haven)
library(readxl)
library(rlang)
library(corrplot)
```

# Read in Data 

```{r readData}

hiit <- read_sav("assignments/BasicTests_HIITMAX.sav")


```


# Question 1

The HIITMAX study was a randomized controlled trial (RCT) run by David comparing the effects of high- versus moderate-intensity on the body composition and physical fitness of 28 overweight young men in the Birmingham area. The measured variables in the dataset have two versions reflecting the longitudinal design: a baseline value ( BL) and a final follow-up value ( F). Because both interventions involved putting previously sedentary subjects onto exercise regimens, David is interested if there was an overall change in respiratory fitness (measured by V O2,max) from baseline to follow-up. Test the investigator’s hypothesis using the data.

## Question/Answer 1a

What test is appropriate for this data and what are the null and alternative hypotheses?

**The appropiate test for this data is a paired t-test.  The null hypotheses is that there is no change in VO_2 max from baseline.  The alternative hypotheses is there is a change from baseline for the VO_2 max.**

## Question/Answer 1b

What are the sample means of V O2,max pre- and post-intervention? 

```{r 1b}

meanBL <- mean(hiit$VO2peak_BL)

meanBL

meanF <- mean(hiit$VO2peak_F)

meanF

```

**The mean pre-interventation is `r paste0(round(meanBL, 2))`.**

**The mean post-intervention is `r paste0(round(meanF, 2))`.**

## Question/Answer 1c

What is the sample mean of the change in V O2,max from baseline to follow-up?

**The sample mean change from baseline to follow is `r round(meanF - meanBL,2)`.**

## Question/Answer 1d

What is the two-tailed p-value associated with that test statistic?

```{r 1d}

q1T <- t.test(hiit$VO2peak_BL, hiit$VO2peak_F, paired = TRUE, alternative = 'two')

print(q1T)

```

**The two-sided p-value for the test statistic is `r q1T$p.value`.** 

## Question/Answer 1e

Write a sentence summarizing this result.

**With a p-value of `r q1T$p.value`, the results indicate that there is a difference between the post and pre treatment measurements.**

# Question 2

The primary purpose of the study was to test for whether the different exercise plans had a different effect on the subjects. That is, is there a ‘difference of differences’ for VO2,max between the HIIT and MIT groups? Test this hypothesis using the data. (Hint: The within-subject change scores for V O2,max were not included in the dataset, but may be calculated using the data step.)

## Question/Answer 2a

What test is appropriate for this data and what are the null and alternative hypotheses?


## Question/Answer 2b

What is the average change in V O2,max for the MIT group? For HIIT?

```{r 2b}

q2 <- hiit %>% 
  mutate(vo2Difference = VO2peak_F-VO2peak_BL) %>% 
  group_by(treatment) %>% 
  select(-ID) %>% 
  summarise(across(.cols = everything(), .fns = list(mean = mean, sd = sd)))

```

The average change in VO_2 max for the MIT group is `r q2$


