---
title: "week7Homework"
author: "M. Onimus"
date: "10/21/2020"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 3
    toc_float:
      collapse: yes
      smooth_scrolling: yes

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r libs, echo=FALSE, warning=FALSE, message=FALSE}
library(ppcor)
library(tidyverse)
library(gtsummary)
library(gt)
library(gtsummary)
library(haven)
library(readxl)
library(rlang)
library(rstatix)
```

# Read in Data

```{r readData}

week7 <- read_sav("assignments/Generalized_phmc.sav") %>% 
  mutate(MENTAL3A = as.factor(MENTAL3A))


```

# Question 1

We are interested in quantifying a possible association between a subject’s BMI and whether or not they have been diagnosed with a mental illness in this population. We also wish to control for age, sex, and race (white, black, and other) to account for some basic confounding.

## Question/Answer 1a

Given the above model, write out the null and alternative hypotheses for a test of the association between mental illness and BMI. 

$log(\frac{\pi}{(1-\pi)}) = \beta_0 + \beta_1BMI + \beta_2Age + \beta_3Sex + \beta_4Race $

$H_0: \beta_1 = 0 $
$H_1: \beta_2 \neq 0 $


## Question/Answer 1b

What is the test statistic for the above test? What is its distribution and associated p-value?


```{r 1b}

lm <- glm(MENTAL3A ~ BMI + RESPAGE + SEX01 + Race, family = 'binomial', data = week7)

# find the odds ratio using exp

summary(lm)

```


**In SAS, the $\chi^2$ = 14.3347 with an associated p-value of 0.0002.**

**In R, I am provided with a z-value of -3.618 with an associated p-value of 0.000296.**

## Question/Answer 1c

What are the null and alternative hypotheses for testing an association between age and kyphosis, using your logistic regression model? 

$H_0:\beta_1 = 0$ 
$H_A:\beta_1 \neq 0$

## Question/Answer 1d

What is the odds ratio for age, along with its 95% CI?

```{r 1d}

lm <- glm(Kyphosis ~ Age, family = 'binomial', data = week7)

# find the odds ratio using exp

odds <- exp(coefficients(lm))

# or use another package to pull it out

oddsEPI <- epiDisplay::logistic.display(lm)

```

**The odds ratio point estimate with the 95% CI is `r paste(oddsEPI$table[[1]])`.**


## Question/Answer 1e

What are the results of inference? Report the test statistic, its distribution, and the associated p-value. 
```{r 1e}

summary(lm)


```

**In SAS, the $\chi^2$ = 0.5393 with an associated p-value of 0.4627.**

**In R, I am provided with a z-value of 0.734 with an associated p-value of 0.463.**


## Question/Answer 1h

Write a brief interpretation of these results.

**Child's age in the sample was not signifanctly associated with the odds of kyphosis (z-value = 0.734, p-value = 0.463).  Additionally, each additional month increased the odds by 0.43% (95% CI: 0.9929, 1.0159).**

