---
title: "week5Homework"
author: "M. Onimus"
date: "10/8/2020"
output:
  html_document: 
    code_folding: hide
    toc: yes
    toc_depth: 3
    toc_float:
      collapse: yes
      smooth_scrolling: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r libs, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(gtsummary)
library(gt)
library(gtsummary)
library(haven)
library(readxl)
library(rlang)
library(rstatix)
```

# Read in Data

```{r readData}

week5 <- read_sav("assignments/GLM_homework_phmc.sav") 

```

# Question 1

David is hoping to come up with a linear model that can be used to predict a person’s BMI using a subset of the PHMC community health survey database. 

We first wish to examine the relationship between BMI and age. 

## Question/Answer 1a

What is the correlation between age and BMI? 

```{r 1a}

lm <- lm(BMI ~ RESPAGE, data = week5)



#lm

```

**The correlation between BMI and age is `r lm$coefficients[[2]]`.**


## Question/Answer 1b

Write out the regression equation for the prediction of BMI using subject age. 

**$BMI = \beta_0 + \beta_1*age$**

## Question/Answer 1c

What are the null and alternative hypotheses for the test for an association between age and BMI, using your model?  

**$H_0:\beta_1 = 0$**
**$H_1:\beta_1\neq0$**

## Question/Answer 1d

What is the estimated regression equation? Also, provide a plot of it against a scatter plot of the data.  

**$BMI = 27.7601 + 0.0043*age$**

```{r 1d}
ggplot(week5, aes(x = RESPAGE, y = BMI)) + 
  geom_point() +
  geom_smooth(method = 'lm') +
  theme_classic()

```

## Question/Answer 1e

What are the results of inference about the association (parameter estimate table)?

```{r 1e}

summary(lm)

```

The results of interference indicate that the two parameters are not signifcantly associated with each other ($t_1=1.075, p = 0.282$) at an $\alpha=0.05$ level.

## Question/Answer 1f

How do the assumptions of linear regression hold up in this case? Use the appropriate plots to support your discussion.  

```{r 1f}

ggplot(data = NULL, aes(x = lm$fitted.values, y = lm$residuals)) +
  geom_point() +
  geom_hline(yintercept = 0) +
  labs(
    x = "Predicted Value",
    y = "Residual"
  ) +
  theme_classic()

```

**The residual/predicted plot above shows a pretty equal distrubution of points above and below 0 although there may be less points below 0.**

```{r }

ggplot(data = NULL, aes(sample = lm$residuals)) +
  geom_qq() +
  theme_classic()

ggplot(data = NULL, aes(x = lm$residuals)) +
  geom_histogram() +
  theme_classic()


```

**The qqplot and the histogram of residuals both show distrubtions which seem to not be normal.  The qqplot looks closer to an exponential plot, not a straight line, and the histrogram shows tailing.**

## Question/Answer 1g

Write a few sentences describing the results. 

**A correlation between BMI and age was investigated using a linear regression.  With a correlation coefficent of 0.0043 and a $p$-value of 0.282, there is no significant correlation between the two variables.  Additionally, the data failed to meet qqplot and histogram normality assumptions** 